---
title: "ML Project"
author: "Cody"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## ML Project

Goal: predict good exercise habits with accelerometer indicators

First question: does each quality level of the exercise have enough data around it?

```{r message=FALSE, echo=FALSE}
#Libraries used for project
library(caret)
library(dplyr)
library(ggplot2)
set.seed(1111)

#Links provided for data in assignment description
linkTrain <- 'https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv'
linkTest <- 'https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv'

#download files to local device from links provided
linkTrain1 <- download.file(url = linkTrain, destfile = "linkTrain.csv", method = "curl")
linkTest1 <- download.file(url = linkTest, destfile = "linkTest.csv", method = "curl")

#read in the downloaded data to R session
trainData <- read.csv('linkTrain.csv')
testData <- read.csv('linkTest.csv')

#turn training data character into numeric factors, have to turn rest of classe sets into factors too
trainData <- transform(trainData, classe = as.numeric(as.factor(trainData$classe)))

#exploratory plot to make sure we have sufficient data for every classe type
hist(trainData$classe)
```
Each quality level does appear to have enough data

## Feature Plot

Looking for some kind of relationship between the range of motion and exercise quality

```{r, echo=FALSE}
#get rid of the near zero variablity factors
trainData <- trainData[-nearZeroVar(trainData)]
testData <- testData[-nearZeroVar(testData)]

#grabbing only the numeric columns so we can do pca, can only do pca on numeric matricies to allow inversion
trainData <- select_if(trainData, is.numeric)
testData <- select_if(testData, is.numeric)

#omitting Na's so we can build a model from the data
trainData <- na.omit(trainData)
testData <- na.omit(testData)

#difference between max and min for range of motion here --
trainData$yawROMarm <- (trainData$max_yaw_arm - trainData$min_yaw_arm)
trainData$pitchROMbelt <- (trainData$max_picth_belt - trainData$min_pitch_belt)
trainData$pitchROMdumbell <- (trainData$max_picth_dumbbell - trainData$min_pitch_dumbbell)
trainData$pitchROMforearm <- (trainData$max_picth_forearm - trainData$min_pitch_forearm)

#exploratory plot to see how max pitch interacts with the quality of movement
featurePlot(x = trainData[,c("yawROMarm","pitchROMbelt", "pitchROMdumbell", "pitchROMforearm")], y = trainData$classe, plot = "pairs")

#training set should have the same columns as the test set in order
# to train with ML using caret library
common_column_names <- intersect(colnames(trainData) , colnames(testData))
common_column_names <- append(common_column_names, 'classe')
trainData <- trainData %>% select(all_of(common_column_names))

```

## Training algorithm

Finding variables in our data set that have a correlation > 0.9 with another variable

```{r, echo=FALSE}
#pca analysis to show which columns are highly correlated
M <- abs(cor(trainData[,-99]))
diag (M) <- 0 
which(M > 0.9, arr.ind=T)
```
We should use Principal Component Analysis to process this data because there are a lot of very strong relationships

## The predicitions vs the actual values, a clear linear relationship is shown

Plotting how well our prediction algorithm did in the training data set

```{r, echo=FALSE}
#Fitting the ML model
modelFit <- train(classe ~ ., method="glm", preProcess="pca", data = trainData)

#create predictions
predTrain <- round(predict(modelFit, subset(trainData, select = -c(classe))))
pred <- round(predict(modelFit, testData))

#limit predictions to the factors that actually show up in dataset 
predTrain[predTrain>5] <- 5
predTrain[predTrain<1] <- 1

pred[pred>5] <- 5
pred[pred<1] <- 1

#plotting the predictions vs the actual to see if there is a linear relationship
plot(predTrain, trainData$classe)
```

## Confusion matrix from the training dataset

Printing the actual confusion matrix information from our training data set

```{r, echo=FALSE}
#comparing against actual values
confusionMatrix(as.factor(trainData$classe), as.factor(predTrain))
```

